Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth

### 2020-11-12 analysis/VOCC/2-biomass-depth-only-model.Rmd
fix to work with dplyr and sdmTMB updates

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    # species: "Redstripe Rockfish"
    # species: "Shortbelly Rockfish"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
    update_predictions: FALSE
    update_model_check: FALSE
---

@@

covars <- readRDS("data/event-covariates.rds")
data <- dplyr::left_join(biomass, covars)

@@

Run climate independent sdmTMB model
```{r }
if (params$update_model) {
  tictoc::tic()
  if (any(names(data) == "adult_density")) {
    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
    ))

@@

   try({ imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      # ar1_fields = TRUE, # changed to FALSE July 2020
      # reml = TRUE,
      include_spatial = TRUE,
      nlminb_loops = 2, 
      newton_steps = 2,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, ".rds"))
   })

  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

  try({
    imm_biomass <- readRDS(paste0(
      "data/", spp,
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))

    nd <- nd_all %>%
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(imm_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions, file = paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
    ))
  })
} else {
  total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "English Sole"
    # species: "Redstripe Rockfish"
    # species: "Shortbelly Rockfish"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
    update_predictions: FALSE
    update_model_check: FALSE
---

@@

covars <- readRDS("data/event-covariates.rds") %>% select(-geometry)
data <- left_join(biomass, covars)

@@

Run climate independent sdmTMB model
```{r }
if (params$update_model) {

  # if updating model, will need new version of mesh to work with changes to sdmTMB
  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots, type = "kmeans")

  tictoc::tic()
  if (any(names(data) == "adult_density")) {
    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
    ))

@@

   # try({ imm_formula <- as.formula(paste(
   #    "imm_density ~ 0 + as.factor(year)", covariates, ""
   #  ))
   # 
   #  imm_biomass <- sdmTMB::sdmTMB(
   #    data = data,
   #    imm_formula,
   #    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
   #    time = "year", spde = spde,
   #    family = tweedie(link = "log"),
   #    # ar1_fields = TRUE, # changed to FALSE July 2020
   #    # reml = TRUE,
   #    include_spatial = TRUE,
   #    nlminb_loops = 2, 
   #    newton_steps = 2,
   #    enable_priors = priors,
   #    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
   #    silent = FALSE
   #  )
   # 
   #  saveRDS(imm_biomass,
   #    file = paste0("data/", spp, "/mod-imm-biomass-", 
   #      spp, covs, "-", ssid_string, ".rds"))
   # })
   #  
  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

  # try({
  #   imm_biomass <- readRDS(paste0(
  #     "data/", spp,
  #     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  #   ))
  # 
  #   nd <- nd_all %>%
  #     filter(ssid %in% model_ssid) %>%
  #     filter(year %in% unique(imm_biomass$data$year))
  #   nd <- na.omit(nd)
  #   nd$year <- as.integer(nd$year)
  # 
  #   imm_predictions <- predict(imm_biomass, newdata = nd)
  #   saveRDS(imm_predictions, file = paste0(
  #     "data/", spp,
  #     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
  #   ))
  # })
} else {
  total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
